Combining information from photometric and spectroscopic data:\\ Don't waste your time in spectrophotometric calibration!

### 2014-07-28 tex/ms.tex
comments in color, and more figure captions, many small changes

% ---- Hogg (aastex) mode ----
\documentclass[12pt, letterpaper, preprint]{aastex}
% ----

% ---- Charlie (emulateapj) mode ----
% \documentclass[iop,numberedappendix]{emulateapj}
% \usepackage{apjfonts}
% ----

\bibliographystyle{apj}
\input{vc}

@@

If you don't believe the reported data noise variances $\sigma_n^2$,
it often makes sense to add in quadrature a ``jitter'' variance $s^2$.
This is a trivial modification of the chi-squared log-likelihood:
\begin{eqnarray}\label{eq:photometricLF}
\ln p(d\given\theta,s^2) &=& -\frac{1}{2}\,\sum_{n=1}^N \left[\frac{[d_n - \mu_n(\theta)]^2}{[\sigma_n^2 + s^2]} + \ln(2\pi\,[\sigma_n^2 + s^2]) \right]
\quad .
\end{eqnarray}
This modification shows the necessity of including the log-variance
term.
This---equation~(\ref{eq:photometricLF})---is in fact the
log-likelihood function we are going to use for the photometric
measurements (BDJ: IS THIS TRUE?  DWH, Currently I'm not using a
jitter term, but I can and will) in what follows.  In this
formulation, if $\theta$ are the parameters of interest, the jitter
variance $s^2$ is a nuisance noise-model parameter.  The Gaussian
Process will be a generalization of this modification.

@@

DRW/BDJ WORDS ABOUT DATA

The data we use to test our method consists of optical spectroscopy of
XX stellar clusters in M31 for which YY-band HST photometric data
exist.  The optical spectra were obtained using the fiber-fed
Hectoscpec intrument on MMT by \citet{caldwell}.  The wavelength range
of the spectra is $3700-7200 \AA$. The instrumental resolution is
$\sim 5\AA$ across the optical range \citep{fabricant13}.  Wavelength
calibration happened. Sky subtraction happened. The data were
approximately spectrophotometrically calibrated using an average
calibration vector for the night constructed from.... (MORE
HERE). Figure \ref{fig:data} shows the spectra before and after
calibration (SHOULD THIS SHOW ALL THE CLUSTERS OR JUST ONE?).  Discuss
S/N, sky subtraction.

\begin{figure}[h!]
\includegraphics[width = \textwidth]{figures/dfig_b192-g242_020.png}
\caption{The calibrated (left) and uncalibrated (right) spectrum for
  \excluster, along with the uncertainties (bottom panel). The
  spectra are not normalized. \label{fig:data}}
\end{figure}

@@

The spectrum of one solar mass of stars that formed at lookback time
$t$ with metallicity $Z$ is given by $S(\lambda, t, Z [, \beta])$.
Here the additional parameters $\beta$ may include details of the
population synthesis model, such as the IMF slope and adjustments to
the stellar spectra or stellar evolutionary tracks. We will assume
$\beta$ to be fixed for the purposes of this exercise, with no loss of
generality (KEEPING IMF SLOPE CONSTANT FOR NOW).  We will also fix the
shape of the attenuation curve, $k_\lambda \equiv
\tau_\lambda/\tau_V$, to the Milky way average of \citet{CCM89} with
$R_V=3.1$ and assume that $d_L$ is known to high precision from
independent methods and can therefore be treated as fixed.  In
practice, $d_L$ is completely degenerate with $m_*$. We calculate
$S(\lambda, t, Z)$ using the Flexible Stellar Population Synthesis
code \citep[FSPS][]{FSPS}.  We use the Padova isochrones
\citep{Padova} and the MILES stellar spectral library \citep{MILES}.

BDJ IMPROVE DISCUSSION OF BROADENING.  The model spectra are smoothed
by convolution with a Gaussian in wavelength space (to model the MMT
line-spread function) and are interpolated to the wavelength points of
the of the observed spectrum while taking into account the
redshift. The instrumental resolution is given by $\sigma$.  Discuss
intrinsic resolution of model spectra (MILES), what to do in
situations where instrumental resolution is higher, lower or
comparable (ouch) to intrinsic velocity broadening.

@@

We estimate the posterior through MCMC sampling after an initial round
of optimization using Powell's method.  The specific sampling
algorithm we use is the affine-invariant ensemble sampling scheme of
\citet{GW}, as implemented in the \texttt{emcee} package \cite{dfm}
Unlike typical Metropolis-Hastings MCMC sampling, the affine-invariant
sampler does not require the specification of step sizes in each
dimension.  ASSESSING SAMPLER CONVERGENCE

Our prior distributions are in all cases top-hat functions, though
more informative prior distributions may be used.  Informative priors
may be especially appropriate for the calibration parameters $\alpha$
if additional information about the instrument is available or for the
parameters of interest $\theta$, for example in the case that there
is knowledge of the metallicity distribution of the population from
which a given cluster is drawn.

\begin{table}[h!]
\caption{List the parameters and  their allowed ranges.}
\end{table}

\subsection{Results}

We first infer the parameters of interest $\theta$ for a calibrated
spectrum of one of the clusters, using the full YY-band photometry. In
Figure \ref{fig:inferred_spectrum} we show the spectrum resulting
from draws of the parameters $[\theta,\alpha]$ from the posterior
distribution.  This figure also shows the calibrated spectrum on which
the priors were conditioned, and several of the individual components
of the model constructed from each draw.  In Figure
\ref{fig:inferred_sed} we also show the residuals between the data
and the draws from the posterior, for both the spectra and the
photometry. In Figure \ref{fig:inferred_params} we show the
marginalized and joint PDFs for the parameters.

@@

BDJ: Finally, we have fun playing with the number of photometric bands
and the S/N.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{figures/sfig_b192-g242_020_cal.png}
\caption{Spectra resulting from draws from the posterior parameter
space conditioned on the calibrated \excluster~ spectrum and
photometry.  Left: For each of 5 draws from the posterior we show from
top to bottom, three components of the model: the mean cluster
spectrum, the calibration vector, and the mean prediction of the
Gaussian process.  Right: From top to bottom, the full model ({\it
cyan}) compared to the observed, calibrated spectrum ({\it magenta}),
the ratio of the observed to the modeled spectrum, and the difference
between the observed spectrum and the modeled spectrum normalized by
the uncertainty.
\label{fig:inferred_spectrum}}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{figures/pfig_b192-g242_020_cal.png}
\caption{The photometric SED resulting \label{fig:inferred_sed}}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{figures/cfig_b192-g242_020.png}
\caption{Show the calibration vector and our inference of it? \label{fig:inferred_calib}}
\end{figure}

@@

\subsection{Dependence on Photometric Data}
In this section we examine how the constraints on physical (and
calibration) parameters vary as a function of the number and type of
photometric points available.

@@

BDJ INSERT WORDS HERE.  We have demonstrated, using a sample of
stellar clusters in M31, that we can combine photometric and
spectroscopic data in a principled way to infer physical properties.
We do not resort to an arbitrary down-weighting or up-weighting of
certain data. Instead, we explicitly model our uncertain knowledge of
the spectral calibration and infer this calibration jointly with the physical
properties. For a number of applications this combination of
spectroscopy and photometry reduces the need for spectrophotometric
calibration.  Calibration information, if it exists, can be
incorporated into informative priors on the calibration parameters of the model.

Our strategy is easily extensible to more complicated models, which
include for example night sky-line emission in the spectroscopy (or
residuals from an intial sky-line subtraction), and uncertainties in
wavelength calibration.  DISCUSS MORE COMPLICATED COVARIANCE KERNELS,
MULTIPLE KERNELS, etc.  MODEL CHECKING.

>>>>

% ---- Hogg (aastex) mode ----
%\documentclass[12pt, letterpaper, preprint]{aastex}
% ----

% ---- Charlie (emulateapj) mode ----
 \documentclass[iop,numberedappendix]{emulateapj}
 \usepackage{apjfonts}
% ----
\usepackage{color}
\bibliographystyle{apj}
\input{vc}

@@

If you don't believe the reported data noise variances $\sigma_n^2$,
it often makes sense to add in quadrature a ``jitter'' variance $s^2$.
This is a trivial modification of the chi-squared log-likelihood:
\begin{eqnarray}\label{eq:photometricLF}
\ln p(d\given\theta,s^2) &=& -\frac{1}{2}\,\sum_{n=1}^N \left[\frac{[d_n - \mu_n(\theta)]^2}{[\sigma_n^2 + s^2]} + \ln(2\pi\,[\sigma_n^2 + s^2]) \right]
\quad .
\end{eqnarray}
This modification shows the necessity of including the log-variance
term.
This---equation~(\ref{eq:photometricLF})---is in fact the
log-likelihood function we are going to use for the photometric
measurements ({\color{blue} BDJ: IS THIS TRUE?} {\color{red} DWH,
Currently I'm not using a jitter term, but I can and will.  Update:
having difficulty with a jitter term}) in what follows.  In this
formulation, if $\theta$ are the parameters of interest, the jitter
variance $s^2$ is a nuisance noise-model parameter.  The Gaussian
Process will be a generalization of this modification.

@@

The data we use to test our method consist of optical spectroscopy of
XX stellar clusters in M31 for which YY-band HST photometric data
exist.  The optical spectra were obtained using the fiber-fed
Hectoscpec intrument on MMT by \citet{caldwell}.  The wavelength range
of the spectra is $3700-9100 \AA$. The instrumental resolution is
$\sim 5\AA$ across the optical range \citep{fabricant13}.  {\color{blue} Wavelength
calibration happened. Sky subtraction happened.} The data were
approximately spectrophotometrically calibrated using an average
calibration vector for the night constructed from {\color{blue} MORE
HERE}. Figure \ref{fig:data} shows the spectra before and after
calibration {\color{blue} SHOULD THIS SHOW ALL THE CLUSTERS OR JUST ONE?.  Discuss
S/N, sky subtraction.}

{\color{red} Specify order of calibration polynomial here?}

\begin{figure*}[h!]
\includegraphics[width = \textwidth]{figures/dfig_b192-g242_020.png}
\caption{The calibrated (left) and uncalibrated (right) spectrum for
  \excluster, along with the uncertainties (bottom panel). The
  spectra are not normalized. \label{fig:data}}
\end{figure*}

@@

The spectrum of one solar mass of stars that formed at lookback time
$t$ with metallicity $Z$ is given by $S(\lambda, t, Z [, \beta])$.
Here the additional parameters $\beta$ may include details of the
population synthesis model, such as the IMF slope and adjustments to
the stellar spectra or stellar evolutionary tracks. We will assume
$\beta$ to be fixed for the purposes of this exercise, with no loss of
generality {\color{blue} (KEEPING IMF SLOPE CONSTANT)}.  We will also fix the
shape of the attenuation curve, $k_\lambda \equiv
\tau_\lambda/\tau_V$, to the Milky way average of \citet{CCM89} with
$R_V=3.1$ and assume that $d_L$ is known to high precision from
independent methods and can therefore be treated as fixed.  In
practice, $d_L$ is completely degenerate with $m_*$. We calculate
$S(\lambda, t, Z)$ using the Flexible Stellar Population Synthesis
code \citep[FSPS][]{FSPS}.  We use the Padova isochrones
\citep{Padova} and the MILES stellar spectral library \citep{MILES}.

The model spectra $F_{\theta^\prime}$ are smoothed by convolution
with a Gaussian in wavelength space (in this case to model the MMT
line-spread function) and are interpolated to the wavelength points of
the of the observed spectrum while taking into account the
redshift. The instrumental resolution is given by the parameter
$\sigma$.  {\color{blue}Discuss intrinsic resolution of model spectra
(MILES), what to do in situations where instrumental resolution is
higher, lower or comparable (ouch) to intrinsic velocity broadening.}

@@

We estimate the posterior through MCMC sampling after an initial round
of optimization using Powell's method.  The specific sampling
algorithm we use is the affine-invariant ensemble sampling scheme of
\citet{GW}, as implemented in the \texttt{emcee} package \cite{dfm}
Unlike typical Metropolis-Hastings MCMC sampling, the affine-invariant
sampler does not require the specification of step sizes in each
dimension.  {\color{blue}ASSESSING SAMPLER CONVERGENCE}

Our prior distributions are in all cases top-hat functions, though
more informative prior distributions may be used.  Informative priors
may be especially appropriate for the calibration parameters $\alpha$
if additional information about the instrument is available; or for the
parameters of interest $\theta$, for example in the case that there
is knowledge of the metallicity distribution of the population from
which a given cluster is drawn.

\input{tables/priors.tbl}
\subsection{Results}

We first infer the parameters of interest $\theta$ for a calibrated
spectrum of one of the clusters, \excluster, using the full YY-band
photometry. In Figure \ref{fig:inferred_spectrum} we show the spectra
that result from draws of the parameters $[\theta,\alpha]$ from the
posterior distribution, as well as several of the individual
components of the model constructed from each draw.  This figure also
shows the calibrated spectrum on which the priors were conditioned,
and residuals between the posterior models and the observed data.  In
Figure \ref{fig:inferred_sed} we also show the residuals between the
data and the draws from the posterior for the spectra. In Figure
\ref{fig:inferred_params} we show the marginalized and joint PDFs for
a subset of the full model parameters.

@@

\begin{figure*}[h!]
\includegraphics[width=\textwidth]{figures/sfig_b192-g242_020_cal.png}
\caption{Spectra resulting from draws from the posterior parameter
space conditioned on the calibrated \excluster~ spectrum and
photometry.  Left: For each of 5 draws from the posterior we show from
top to bottom, three components of the model: the mean cluster
spectrum, the calibration vector, and the mean prediction of $\Delta$
for the Gaussian process.  Right: From top to bottom, the full model
({\it cyan}) compared to the observed, calibrated spectrum ({\it
magenta}), the ratio of the observed to the modeled spectrum, and the
difference between the observed spectrum and the modeled spectrum
normalized by the uncertainty.
\label{fig:inferred_spectrum}}
\end{figure*}

\begin{figure}[h!]
\includegraphics[width=0.5\textwidth]{figures/pfig_b192-g242_020_cal.png}
\caption{Model SEDs.  Top: The model photometric SED resulting from draws from the posterior
  parameter space conditioned ({\it cyan}) on the observed photometry
  ({\it magenta})  and spectra.  Bottom: uncertainty normalized
  residuals of the models from the observations. 
 . \label{fig:inferred_sed}}
\end{figure}

@@

\begin{figure*}[h!]
\includegraphics[width=\textwidth]{figures/cfig_b192-g242_020.png}
\caption{Show the calibration vector and our inference of it? \label{fig:inferred_calib}}
\end{figure*}

@@

\subsection{Dependence on Photometric Data}
In this section we examine how the constraints on physical (and
calibration) parameters vary as a function of the number and type of
photometric points available. {\color{red} BDJ DO THIS TEST.}

@@

We have demonstrated, using a sample of stellar clusters in M31, that
we can combine photometric and spectroscopic data in a principled way
to infer physical properties.  We do not resort to an arbitrary
down-weighting or up-weighting of certain data. Instead, we explicitly
model our uncertain knowledge of the spectral calibration and infer
this calibration jointly with the physical properties. For a number of
applications this combination of spectroscopy and photometry reduces
the need for spectrophotometric calibration.  Calibration information,
if it exists, can be incorporated into informative priors on the
calibration parameters of the model.

Our strategy is easily extensible to more complicated models, which
include for example night sky-line emission in the spectroscopy (or
residuals from an intial sky-line subtraction), and uncertainties in
wavelength calibration. {\color{blue} DISCUSS MORE COMPLICATED
COVARIANCE KERNELS, MULTIPLE KERNELS, etc.  MODEL CHECKING.}
